Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-06 ms.tex
more challenges to quell

\caption{Summary of Numerical Solutions}

@@

However, even if SN Ia are rare near the stagnation radius, SN Ia may cap the
SMBH accretion rate by periodically blowing gas out of the nucleus of
low mass galaxies.  Between successive SNe, stars release a
gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$
interior to the Ia radius, which is locally gravitationally bound to the SMBH
by an energy $E_{\rm bind} \sim M_{\rm g}\sigma_0^{2}\mathcal{F}^{2}$, where typically $\mathcal{F} 
\lesssim$ few.  From the above definitions we find that 
\be
\frac{E_{\rm Ia}}{E_{\rm bind}} \sim \frac{(v_{w}^{\rm Ia})^{2}}{2\mathcal{F}^{2}\sigma_0^{2}}.  
\label{eq:blowout}
\ee
Hence, for low mass BHs with $\sigma_0 \ll \mathcal{F}v^{\rm Ia}_{w}/2$, SN Ia are capable of dynamically clearing out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Even when heating is sufficiently weak that the stagnation radius formally exceeds $r_{\rm Ia}$, the SMBH accretion rate is still potentially limited to a value
\begin{eqnarray}
\frac{\dot{M}_{\rm Ia}}{\dot{M}_{\rm edd}} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{r_{\rm Ia}}{r_{\rm soi}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    4.5 \times 10^{-4} M_{\bullet,8}^{-1.33}(\tau_{\star}/t_{\rm h})^{-0.2}
   & \text{core} \\
    2.2 \times 10^{-4} \Mbheight^{-0.84}(\tau_{\star}/t_{\rm h})^{-0.6}   & \text{cusp},
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
obtained substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}]) for $r_{\rm s}$ in the derivation leading to our estimate of $\dot{M}$ (eq.~[$\ref{eq:eddr_analytic}$]).  Note, however, that for low mass SMBHs, $\dot{M}_{\rm Ia}$ generally exceeds the maximum accretion rate for a thermally stable flow $\dot{M}_{\rm TI}$ (eq.~[\ref{eq:Mdotmax}]), implying that dynamical `blow-out' from SNe Ia cannot alone prevent cooling instabilities in the nuclei of low mass galaxies.

@@

For low-mass black holes (Fig.~\ref{fig:vwSourcesImp}, top panel), stellar winds are the most important source
of heating on timescales of $\tau_{\star} \lesssim 10^{7}$
years.  Compton heating comes to dominate at later times due to the
higher accretion rates that accompany the overall decrease in the
heating rate, coupled with the persistently high mass loss rates and
gas densities associated with a relatively young stellar population.  Interestingly, Ia SN dominate the heating briefly around 10$^{9}$ yr due to a window of time in which $r_Ia < \rs$.  For high-mass black holes (Fig.~\ref{fig:vwSourcesImp}, bottom panel), Compton heating dominates at all epochs, while SN Ia  dominate after 10$^{8}$ yr.

@@

Figure \ref{fig:bh_growth} (top panel) shows with solid lines the
accretion rate as a function of black hole mass for the average star
formation heating, calculated from equation (\ref{eq:eddr_analytic})
using our results for the total wind heating
(Fig.~\ref{fig:mdot_mass}).  This average accretion rate increases
from $\dot{M} \sim 10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ for low-mass
black holes ($M_{\bullet} \sim 10^{6}M_{\odot}$) to $\dot{M} \sim
10^{-4}M_{\odot}$ yr$^{-1}$ for $M_{\bullet} \sim 10^{9}M_{\sun}$.
These fall below the maximum thermally stable accretion rate,
$\dot{M}_{\rm TI}$ (yellow lines), although they become closer at the highest
black hole masses $\sim 10^{9}M_{\odot}$.

@@

    \begin{enumerate}
    \item Overall the accretion rate onto the central SMBH will be strongly
    tied to the properties of the stellar population (e.g. age) within
    the sphere of influence of the black hole. Thus, our model cannot
    say anything definitive about the mode accretion for individual
    galaxies which may have significant variations about our assumed
    average star formation histories.
    \item Also, the star formation history of bulges may (and probably
      does) systematically differ from the average star formation
      history for galaxies as a whole.
    \item Furthermore, there could be a radial gradient in the star
      formation history within a given bulge... 
    \end{enumerate}

>>>>

  \caption{Summary of Numerical Solutions {\bf AG make sure the
      headings are consistent. Mark solutions which are unstable for
      some values of eta.}}

@@

However, even if SN Ia are rare near the stagnation radius, SN Ia may cap the
SMBH accretion rate by periodically blowing gas out of the nucleus of
low mass galaxies.  Between successive SNe, stars release a
gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$
interior to the Ia radius, which is locally gravitationally bound to the SMBH
by an energy $E_{\rm bind} \sim M_{\rm g}\sigma_0^{2}\mathcal{F}^{2}$, where typically $\mathcal{F} 
\lesssim$ few.  From the above definitions we find that 
\be
\frac{E_{\rm Ia}}{E_{\rm bind}} \sim \frac{(v_{w}^{\rm Ia})^{2}}{2\mathcal{F}^{2}\sigma_0^{2}}.  
\label{eq:blowout}
\ee
Hence, for low mass BHs with $\sigma_0 \ll \mathcal{F}v^{\rm Ia}_{w}/2$, SN Ia are capable of dynamically clearing out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Even when heating is sufficiently weak that the stagnation radius formally exceeds $r_{\rm Ia}$, the SMBH accretion rate is still potentially limited to a value
\begin{eqnarray}
\frac{\dot{M}_{\rm Ia}}{\dot{M}_{\rm edd}} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{r_{\rm Ia}}{r_{\rm soi}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    4.5 \times 10^{-4} M_{\bullet,8}^{-1.33}(\tau_{\star}/t_{\rm h})^{-0.2}
   & \text{core} \\
    2.2 \times 10^{-4} \Mbheight^{-0.84}(\tau_{\star}/t_{\rm h})^{-0.6}   & \text{cusp},
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
obtained substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}])
for $r_{\rm s}$ in the derivation leading to our estimate of $\dot{M}$
(eq.~[$\ref{eq:eddr_analytic}$]).  Note, however, that for low mass
SMBHs, $\dot{M}_{\rm Ia}$ generally exceeds the maximum accretion rate
for a thermally stable flow $\dot{M}_{\rm TI}$
(eq.~[\ref{eq:Mdotmax}]), implying that dynamical `blow-out' from SNe
Ia cannot alone prevent cooling instabilities in the nuclei of low
mass galaxies {\bf AG--careful one needs a separate criterion for
  thermal stability when heating is dominated by the Ias}.

@@

For low-mass black holes (Fig.~\ref{fig:vwSourcesImp}, top panel),
stellar winds are the most important source of heating on timescales
of $\tau_{\star} \lesssim 10^{7}$ years.  Compton heating comes to
dominate at later times due to the higher accretion rates that
accompany the overall decrease in the heating rate, coupled with the
persistently high mass loss rates and gas densities associated with a
relatively young stellar population.  Interestingly, Ia SN dominate
the heating briefly around 10$^{9}$ yr due to a window of time in
which $r_{ \rm Ia} < \rs$.  For high-mass black holes
(Fig.~\ref{fig:vwSourcesImp}, bottom panel), Compton heating dominates
at all epochs, while SN Ia dominate after 10$^{8}$ yr.

@@

Figure \ref{fig:bh_growth} (top panel) shows with solid lines the
accretion rate as a function of black hole mass for the average star
formation heating, calculated from equation (\ref{eq:eddr_analytic})
using our results for the total wind heating
(Fig.~\ref{fig:mdot_mass}).  This average accretion rate increases
from $\dot{M} \sim 10^{-6}-10^{-5}M_{\odot}$ yr$^{-1}$ for low-mass
black holes ($M_{\bullet} \sim 10^{6}M_{\odot}$) to $\dot{M} \sim
10^{-4}M_{\odot}$ yr$^{-1}$ for $M_{\bullet} \sim 10^{8}M_{\sun}$.
These fall below the maximum thermally stable accretion rate,
$\dot{M}_{\rm TI}$ (yellow lines).
% , although they become closer at the highest
% black hole masses $\sim 10^{9}M_{\odot}$.

@@

    \begin{enumerate}
    \item Overall the accretion rate onto the central SMBH will be strongly
    tied to the properties of the stellar population (e.g. age) within
    the sphere of influence of the black hole. Thus, our model cannot
    say anything definitive about the mode accretion for individual
    galaxies which may have significant variations about our assumed
    average star formation histories.
    \item Also, the star formation history of bulges may (and probably
      does) systematically differ from the average star formation
      history for galaxies as a whole.
    \item Furthermore, there could be a radial gradient in the star
      formation history within a given bulge... 
     {\bf AG this section needs to be finished...}
    \end{enumerate}
